---
title: "ReproResProj1"
author: "Kevin Goldsmith"
date: "3/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible Research Project 1 

####1) Code for reading in the dataset and/or processing the data  
Load the data 
```{r}
fitdata<-read.csv("activity.csv")
```

Process/transform the data into a suitable format for analysis 
```{r warning= FALSE}
fitdata$date<-as.Date(as.character(fitdata$date), 
                   format = "%Y-%m-%d")
fitdataNB<-na.omit(fitdata)
```

####2) Histogram of the total number of steps taken each day  
Calculate the total number of steps taken per day
```{r}
library(plyr)
daydata<- ddply(fitdataNB,.(date),summarize,TotalSteps=sum(steps))
```

Make a histogram of the total number of steps taken each day
```{r}
hist(daydata$TotalSteps, main = "Histogram of steps per day", xlab = "Total Steps per Day")
```

####3) Mean and median number of steps taken each day
Calculate and report the mean and median of the total number of steps taken per day

```{r}
meansteps = mean(daydata$TotalSteps)
mediansteps = median(daydata$TotalSteps)
cat("Mean: ",  meansteps, "\nMedian: ", mediansteps)
```

####4) Time series plot of the average number of steps taken
Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days 
```{r}
plot(daydata, type = "l", main = "Total Steps by Day")
```

####5)The 5-minute interval that, on average, contains the maximum number of steps
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
intervalavg<-ddply(fitdataNB, .(interval), summarize, AvgSteps = mean(steps))
maxsteps = which.max(intervalavg$AvgSteps)
cat("Higest average interval: ", intervalavg$interval[maxsteps])
```

####6) Code to describe and show a strategy for imputing missing data
Calculate and report the total number of missing values in the dataset 
```{r}
NAcount = sum(is.na(fitdata$steps))
print(NAcount)
```

Devise a strategy for filling in all of the missing values in the dataset. Use the mean for the 5-minute interval. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r message= FALSE, warning = FALSE}
library(dplyr)
fitdata2<-fitdata %>% group_by(interval) %>% mutate(steps = ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps)) %>% ungroup
daydata2<- ddply(fitdata2,.(date),summarize,TotalSteps=sum(steps))
```

####7) Histogram of the total number of steps taken each day after missing values are imputed
Make a histogram of the total number of steps taken each day
```{r}
hist(daydata2$TotalSteps, main = "Histogram of steps per day", xlab = "Total Steps per Day")
```

Calculate and report the mean and median total number of steps taken per day

```{r}
meansteps = mean(daydata2$TotalSteps)
mediansteps = median(daydata2$TotalSteps)
cat("Mean: ",  meansteps, "\nMedian: ", mediansteps)
```

####8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
Create a new factor variable in the dataset with two levels - "weekday" and "weekend"
```{r}
byweekday<-mutate(fitdata2,WD = factor(ifelse(weekdays(fitdata2$date) %in% c("Saturday", "Sunday"),"Weekend","Weekday")))
```

Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged acrossed all weekdays or weekend days 

```{r}
library(lattice)
intervaldata<- ddply(byweekday,c(.(interval),.(WD)),summarize,AvgSteps=mean(steps))
xyplot(AvgSteps ~ interval | WD, data = intervaldata, layout = c(1,2), type = "l")
```


